---
title: "INFO634 - Group project"
author: "Below analisys has been done by Nadiia Nikoniuk"
date: "11/09/2021"
output:
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
---
## Loading vaccinations data from STATS NZ using API

### Explanatory data analysis of loaded datasets

To load the data we have used API published by the Stats NZ.
Dataframe Observations_CPCOV10 contains the cumulative vaccination data around NZ.
Dataframe Observations_CPCOV11 contains the cumulative vaccination data by vaccinator.
```{r setup}
source("get-opendata-catalogue-fun.R")

Catalogue <- get_odata_catalogue(
  service="https://api.stats.govt.nz/opendata/v1",
  endpoint="data.json",
  service_api_key = "3727dfa9d9f44874a510f426cd115b1c"
)


source("get-odata-fun.R")
Observations_CPCOV10<-  Filter(function(x)!all(is.na(x)),
                               get_odata(
                                 service = "https://api.stats.govt.nz/opendata/v1",
                                 endpoint = "Covid-19Indicators",
                                 entity = "Observations",
                                 query_option = "$filter=(
                                                    ResourceID eq 'CPCOV10' 
                                                  )   

                                          ",
                                 service_api_key = "3727dfa9d9f44874a510f426cd115b1c"))
Observations_CPCOV11<-  Filter(function(x)!all(is.na(x)),
                               get_odata(
                                 service = "https://api.stats.govt.nz/opendata/v1",
                                 endpoint = "Covid-19Indicators",
                                 entity = "Observations",
                                 query_option = "$filter=(
                                                    ResourceID eq 'CPCOV11' 
                                                  )   

                                          ",
                                 service_api_key = "3727dfa9d9f44874a510f426cd115b1c"))

#Change Period to Date variable
library(lubridate)
library(DT)
library(tidyverse)
Observations_CPCOV10$Period<-as.Date(Observations_CPCOV10$Period, format="%Y-%m-%d")
Observations_CPCOV11$Period<-as.Date(Observations_CPCOV11$Period, format="%Y-%m-%d")
Observations_CPCOV10$Period_Num<-as.numeric(Observations_CPCOV10$Period)
Observations_CPCOV11$Period_Num<-as.numeric(Observations_CPCOV11$Period)

#Change Label1 variable to factor
Observations_CPCOV10$Label1<-factor(Observations_CPCOV10$Label1)
Observations_CPCOV11$Label1<-factor(Observations_CPCOV11$Label1)
```
Observations_CPCOV10 dataframe contains 84 observations and 9 variables. After the data is loaded there are three data types detected: character, numeric and integer. I have updated the Period to date variable and Label to factor and its cardinality is 3.

Observations_CPCOV11 dataframe contains 56 observations and 11 variables. After the data is loaded there are three data types detected: character, numeric and integer. I have updated the Period to date variable and Label to factor and its cardinality is 2.


## Cumulative vaccination dataset analisys

### Cumulative vaccination: Data overview

Plot the dataset using *ggplot* function

```{r plot}
library(ggplot2)
ggplot(Observations_CPCOV10, aes(x=Period, y = Value, group=Label1, colour=Label1)) + 
  geom_line() + 
  geom_point(size = 1)+
  ggtitle("Cumulative vaccinations")+
  xlab("Week ending") +
  ylab("Number of vaccinations administered")+
  labs(fill = "1")+
  theme(legend.position="top")

```

There are some inconsistency in Cumulative vaccine stock data between June and August.

Data table


```{r cumulative_data_table}
Observations_CPCOV10
```
Summary table

```{r summary}
summary(Observations_CPCOV10)
```
### Cumulative vaccination: Explore dataset for a missing data

For a dataset, it is important to be able to answer these questions about missing values.

```{r missing_cumulative}
library(visdat)
library(ggplot2)
vis_dat(Observations_CPCOV10) + ggtitle("Missing Value Distribution")
```

As per above there is no missing data detected in the dataset.

### Cumulative vaccination: Novelties
```{r outliers1}

ggplot(data = Observations_CPCOV10, aes(x="Label", y=Value)) + geom_boxplot(aes(fill=Label1), outlier.colour = "red")+ 
labs(title = paste("Boxplots: Cumulative vaccinations"))+
  theme(axis.title.y = element_blank(), axis.text.y = element_text(), axis.ticks.y = element_blank())
```

Box plot allows to identify outliers for a numeric data. There are some outliers detected but because this is a cumulative data the latest observations are treated as ouliers.

### Cumulative vaccination: Split data by testing and trainign dataset

To valuate the mothod performance we should split the dataset by testing and training. 
The first 80% of data is assigned to training dataset and the last 20% is assigned to testing dataset.
We will traing our method using training dataset and find the optimal hyper parameters to get the model. This model will be applied to the training dataset. 

```{r sampling}
#separate "Cumulative vaccinations administered" observations as we are going to do a projection just for "Cumulative vaccinations administered"
vaccin_adm<-Observations_CPCOV10[which(Observations_CPCOV10$Label1=="Cumulative vaccinations administered"),] #
#split the data by training and testin
d<-1:(nrow(vaccin_adm)*0.8)
train<-vaccin_adm[d,]
test<-vaccin_adm[-d,]

#order the dataframe by date
train<-train[order(train$Period),]
test<-test[order(test$Period),]
```

### Cumulative vaccination: Vaccination forecast using SVM method 
The first method we selected is SVM. Lets check how it performs.
```{r svm}
library(e1071)
library(Metrics)

#train with hyperparameter tuning
tuneResultsvm <- tune(svm, Value ~ Period,  data = train, kernel = "polynomial",
              ranges = list(gamma=2^(-1:1), epsilon = seq(0,1,0.1), cost = 2^(seq(0.5,10,.5))))
summary(tuneResultsvm$best.model)

#best model selection
tunedVals <-tuneResultsvm$best.model

#best parameters selection
tunePar<-tuneResultsvm$best.parameters
gammaBest<-tunePar$gamma
epsilonBest<-tunePar$epsilon
costBest<-tunePar$cost

#testing using the best model
predictYsvm <- predict(tunedVals, test)

```
### Cumulative vaccination: SVM model performance
```{r errors_svm}
library(Metrics)
#errors
RMSE_svm<-rmse(test$Value, predictYsvm)
MAE_svm<-mae(test$Value, predictYsvm)
RSE_svm<-1-rse(test$Value, predictYsvm)

#creating error table
Measurement<-c("RMSE", "R-squared", "MAE")
err_Value<-c(RMSE_svm, RSE_svm, MAE_svm)
employ.data <- data.frame(Measurement, err_Value)
employ.data
```
Above table is sowing the SVM performs too good and there is a high risk of overfitting. 
We can plot the predicted and testing values to see how our model behaves.

```{r plot_test_svm}


library(ggplot2)
ggplot(test, aes(x=Period, y = Value, group=Label1, colour=Label1)) + 
  geom_line() +
  geom_point(aes(x=test$Period, y=predictYsvm),colour='blue')+
  ggtitle("Cumulative vaccinations administered vs predicted values- SVM")+
  xlab("Week ending") +
  ylab("Number of vaccinations administered")+
  labs(fill = "1")+
  theme(legend.position="top")
```

As per above plot the predicted values are higher then actual at the end which means the projection will overestimate the number of vaccinations.

We should also check the residual outliers for our model. Residual=forecast-test or train

Checking the residuals outliers for training dataset:
```{r plot_test_svm2}


  d_svm <- data.frame(train$Value, tuneResultsvm$best.model$fitted)
  colnames(d_svm) <- c("obs", "pred")
    d_svm$residuals <- tuneResultsvm$best.model$residuals
    limits <- boxplot.stats(x = d_svm$residuals)$stats
    label <- ifelse(d_svm$residuals < limits[1] | d_svm$residuals > limits[5], rownames(d_svm), NA)
  ggplot(d_svm, mapping = aes(y = residuals, x = 0, label = label)) +
      geom_boxplot(orientation = "vertical") +
      labs(title = "Train-Residual Boxplot - SVM") +
      theme(axis.title.x = element_blank(), axis.text.x = element_text(), axis.ticks.x = element_blank())
  
```

There are no residuals outliers detected in a training dataset.

Checking the residuals outliers for testing dataset:

```{r plot_test_svm3}
d_svm_test <- data.frame(test$Value, predictYsvm)
 colnames(d_svm_test) <- c("obs", "pred")
   range <- range(c(d_svm_test$obs, d_svm_test$pred), na.rm = TRUE)
   
   d_svm_test$residuals <- d_svm_test$obs - d_svm_test$pred
    limits <- boxplot.stats(x = d_svm_test$residuals)$stats
    label <- ifelse(d_svm_test$residuals < limits[1] | d_svm_test$residuals > limits[5], rownames(d_svm_test), NA)
   ggplot(d_svm_test, mapping = aes(y = residuals, x = 0, label = label)) +
      geom_boxplot(orientation = "vertical") +
      labs(title = "Test-Residual Boxplot - SVM") +
      theme(axis.title.x = element_blank(), axis.text.x = element_text(), axis.ticks.x = element_blank())
    
```

There are no residuals outliers detected in a testing dataset.

### Cumulative vaccination: Vaccination forecast - SVM
Vaccination projection using SVM.
We have to create a model using the whole dataset and applying the hyperparameters getting above.

```{r future_prediction}

#creating the dataframe for projection 
vaccine_prediction_svm <- data.frame(id=1, ResourceID="CPC0V10", Period = c(    # Create example data
                            
                             "2021-09-12",
                             "2021-09-19",
                             "2021-09-26",
                             "2021-10-03",
                             "2021-10-10",
                             "2021-10-17",
                             "2021-10-24",
                             "2021-10-31",
                             "2021-11-07",
                             "2021-11-14",
                             "2021-11-21",
                             "2021-11-28",
                             "2021-12-05",
                             "2021-12-12",
                             "2021-12-19",
                             "2021-12-26",
                             "2022-01-02",
                             "2022-01-09",
                             "2022-01-16",
                             "2022-01-23",
                             "2022-01-30",
                             "2022-02-06",
                             "2022-02-13",
                             "2022-02-20",
                             "2022-02-27",
                             "2022-03-06",
                             "2022-03-13",
                             "2022-03-20",
                             "2022-03-27"
                             ),
                   Label1 = "Cumulative vaccinations forecast", Value=0, Unit="Number", Measure="Cumulative Vaccinations", Multiplier=0)
vaccine_prediction_svm$Period <- as.Date(vaccine_prediction_svm$Period,             # Change class of date column
                      format = "%Y-%m-%d")
vaccine_prediction_svm$Period_Num<-as.numeric(vaccine_prediction_svm$Period)
vaccine_prediction_svm<-vaccine_prediction_svm[order(vaccine_prediction_svm$Period),]


#creating the model using the whole dataset
svm_model <- svm(Value ~ Period , gamma=gammaBest, epsilon=epsilonBest, cost=costBest, kernel = "polynomial", vaccin_adm)

#projection
vaccine_prediction_svm$Value <- predict(svm_model, vaccine_prediction_svm)

#plot the data
library(ggplot2)
library("ggrepel")
ggplot()+
  geom_point(data=vaccin_adm, aes(x=Period, y = Value, group=Label1, colour=Label1)) + 
  geom_point(data=vaccine_prediction_svm, aes(x=Period, y = Value, group=Label1, colour=Label1))+
  geom_hline(yintercept=7171640, linetype="dashed", color = "blue")+
  annotate("text",x = as.Date("2021-04-07"), y = 7171640, label = "70% population", color = "blue")+
  geom_hline(yintercept=8196160, linetype="dashed", color = "green")+
  annotate("text",x = as.Date("2021-04-07"), y = 8196160, label = "80% population", color = "green")+
  geom_hline(yintercept=9220680, linetype="dashed", color = "red")+
  annotate("text",x = as.Date("2021-04-07"), y = 9220680, label = "90% population", color = "red")+
  ylim(0, 11000000)+
  ggtitle("Cumulative vaccinations forecast - SVM")+
  xlab("Week ending") +
  ylab("Number of vaccinations administered")+
  labs(fill = "1")+
  theme(legend.position="top")

```

The NZ population is 5,122,600 people which means to vaccinate 100% population the  10,245,200 vaccines should be administered. 
As per our projection using SVM: 
-70% population will be vaccinated at the end of October;
-80% population will be vaccinated in November;
-90% population will be vaccinated in December.

As I predicted model created by SVM overestimate vaccination as it doesn't look plausible. I decided to try Polynomial Regression as alternative method to do a vaccination projection.

### Cumulative vaccination: Vaccination forecast using Polynomial Regression method  

```{r lm}
#train model
Result_lm <- lm(Value ~ poly(Period_Num,2), data=train)
summary(Result_lm)

#predict using testing dataset
predictYlm <- predict(Result_lm, test)

```

As per above summary Polynomial Regression is showing a good performance on training dataset. We can check how it performs on testing dataset.

### Cumulative vaccination:Polynomial Regression model performance
```{r lm_errors}

#estimate errors
RMSE_lm<-rmse(test$Value, predictYlm)
MAE_lm<-mae(test$Value, predictYlm)
RSE_lm<-1-rse(test$Value, predictYlm)

#creating error table
Measurement<-c("RMSE", "R-squared", "MAE")
err_Value<-c(RMSE_lm, RSE_lm, MAE_lm)
employ.data <- data.frame(Measurement, err_Value)
employ.data
```
Polynomial Regression model is showing s good results on the testing dataset. Also we can control the function degree adding or removing variables. The optimal number is 2 as it gives us R-squared 0.62, so our model is not overfitting.
We can plot the predicted and testing values to see how our model behaves. 

```{r plot_prediction_lm}


library(ggplot2)
ggplot(test, aes(x=Period, y = Value, group=Label1, colour=Label1)) + 
  geom_line() +
  geom_point(aes(x=test$Period, y=predictYlm),colour='blue')+
  ggtitle("Cumulative vaccinations Actual vs Predicted - Polynomial Regression")+
  xlab("Week ending") +
  ylab("Number of vaccinations administered")+
  labs(fill = "1")+
  theme(legend.position="top")

```

We should also check the residual outliers for our model. Residual=forecast-test or train

Checking the residuals outliers for training dataset:
```{r plot_test_lm2}

  d_lm <- data.frame(train$Value, Result_lm$model$Value)
  colnames(d_lm) <- c("obs", "pred")
    d_lm$residuals <- Result_lm$residuals
    limits <- boxplot.stats(x = d_lm$residuals)$stats
    label <- ifelse(d_lm$residuals < limits[1] | d_lm$residuals > limits[5], rownames(d_lm), NA)
  ggplot(d_lm, mapping = aes(y = residuals, x = 0, label = label)) +
      geom_boxplot(orientation = "vertical") +
      labs(title = "Train-Residual Boxplot - Polynomial Regression") +
      theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank())
  
```

There are no residual outliers are detected in training dataset. 

Checking the residuals outliers for testing dataset:
```{r plot_test_lm3}
d_lm_test <- data.frame(test$Value, predictYlm)
    colnames(d_lm_test) <- c("obs", "pred")
   range <- range(c(d_lm_test$obs, d_lm_test$pred), na.rm = TRUE)
   
   d_lm_test$residuals <- d_lm_test$obs - d_lm_test$pred
    limits <- boxplot.stats(x = d_lm_test$residuals)$stats
    label <- ifelse(d_lm_test$residuals < limits[1] | d_lm_test$residuals > limits[5], rownames(d_lm_test), NA)
   ggplot(d_lm_test, mapping = aes(y = residuals, x = 0, label = label)) +
      geom_boxplot(orientation = "vertical") +
      labs(title = "Test-Residual Boxplot - Polynomial Regression") +
      theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank())
    
```

There are no residual outliers are detected in testing dataset.


### Cumulative vaccination: Vaccination forecast - lm

```{r future_prediction_lm}

#creating the dataset for forecasting
vaccine_prediction_lm<-vaccine_prediction_svm
#creating the model
Result_lm_ <- lm(Value ~ poly(Period_Num,2), data=vaccin_adm)
#forecasting
vaccine_prediction_lm$Value <- predict(Result_lm_, vaccine_prediction_lm)

#plot the data
library(ggplot2)
ggplot()+
  geom_point(data=vaccin_adm, aes(x=Period, y = Value, group=Label1, colour=Label1)) + 
  geom_point(data=vaccine_prediction_lm, aes(x=Period, y = Value, group=Label1, colour=Label1))+
  geom_hline(yintercept=7171640, linetype="dashed", color = "blue")+
  annotate("text",x = as.Date("2021-04-07"), y = 7171640, label = "70% population", color = "blue")+
  geom_hline(yintercept=8196160, linetype="dashed", color = "green")+
  annotate("text",x = as.Date("2021-04-07"), y = 8196160, label = "80% population", color = "green")+
  geom_hline(yintercept=9220680, linetype="dashed", color = "red")+
  annotate("text",x = as.Date("2021-04-07"), y = 9220680, label = "90% population", color = "red")+
  ggtitle("Cumulative vaccinations forecast - Polynomial Regression")+
  ylim(0, 11000000)+
  xlab("Week ending") +
  ylab("Number of vaccinations administered")+
  labs(fill = "1")+
  theme(legend.position="top")

```

As per our projection using SVM: 
-70% population will be vaccinated early December;
-80% population will be vaccinated in the missle of December;
-90% population will be vaccinated in January.

I think Polynomial Regression performs better than SVM as we have a model overfitting using SVM method.


### Cumulative vaccination by vaccinators: Explore the dataset for a missing data

For a dataset, it is important to be able to answer these questions about missing values.

```{r missing1}
library(visdat)
library(ggplot2)
vis_dat(Observations_CPCOV11) + ggtitle("Missing Value Distribution")
```

As per above there is no missing data in the dataset.


### Cumulative vaccination by vaccinators: Plot the data
Belos is a plot which is showing the vaccinations by trained or active vaccinator. 

```{r plot_vaccin}

  ggplot(Observations_CPCOV11, aes(Period, Value, fill = Label1)) +
  geom_col() +
  ggtitle("Trained vaccinators and active vaccinators")+
  xlab("Week ending") +
  ylab("Number of vaccinations administered")+
  labs(fill = "Vaccinator:")+
  theme(legend.position="top")
```

Data table

```{r cumulative_data_table1}
Observations_CPCOV11
```


## Cumulative vaccination by vaccinators: Summary table
```{r summary1}
summary(Observations_CPCOV11)
```

### Cumulative vaccination by vaccinators: Explore dataset for outliers

```{r outliers}
coef1 <- 1.5  
limits <- boxplot.stats(x = Observations_CPCOV11$Value, coef = coef1)$stats
Observations_CPCOV11$label <- ifelse(Observations_CPCOV11$Value < limits[1] | Observations_CPCOV11$Value > limits[5], rownames(Observations_CPCOV11), NA)



ggplot(data = Observations_CPCOV11, aes(x="Label1", y=Value)) + geom_boxplot(aes(fill=Label1), coef = coef1, outlier.colour = "red")+
  labs(title = paste("Uni-variable boxplots at IQR multiplier of", coef1))+
  theme(axis.title.y = element_blank(), axis.text.y = element_text(), axis.ticks.y = element_blank())
```

There are no outliers detected. 

## DHB of Residence by ethnicity: Loading vaccinations data from the Excel file published by the Ministry of Health
To check how NZ performs with vaccination we have downloaded the dataset published by the Ministry of Health. The dataset is dated from 7 September 2021.

Vaccin_DHB dataframe contains 28 observations and 9 variables. After the data is loaded there are three data types detected: character, numeric and integer. I have updated the `DHB of residence`, `Age group`, `Ethnic group`, `Gender` to factor as these variables have a low cardinality.

```{r read_excel1}

library("readxl")
Vaccin_DHB <- read_excel("covid_vaccinations.xlsx", sheet = "DHBofResidence by ethnicity")
#drop last two columns
Vaccin_DHB<-Vaccin_DHB[-c(10,11)]
Vaccin_DHB$`DHB of residence`<-factor(Vaccin_DHB$`DHB of residence`)
Vaccin_DHB$`Age group`<-factor(Vaccin_DHB$`Age group`)
Vaccin_DHB$`Ethnic group`<-factor(Vaccin_DHB$`Ethnic group`)
Vaccin_DHB$`Gender`<-factor(Vaccin_DHB$`Gender`)
Vaccin_DHB
```
Datafreame summary
```{r summarytools}
library(summarytools)
summarytools::view(summarytools::dfSummary(Vaccin_DHB), method = "render")
```

### DHB of Residence by ethnicity: Homogeniety

I have selected numeric variables to check the homogeniety. As per below the data belong together and there is no significant differences in the data.


```{r homogeniety}
cols <- c("First dose administered","Second dose administered","Population","First dose uptake per 1000 people","Second dose uptake per 1000 people") # choose the numeric columns
numData <- scale(Vaccin_DHB[,cols], center = TRUE, scale = TRUE) 
matplot(numData, type = "l", col = rainbow(ncol(numData)), xlab = "Observations in sequence", ylab = "Value") 
```


### DHB of Residence by ethnicity: : Explore dataset for a missing data

Plot the missing ov=bservations.

```{r missing_}
library(visdat)
library(ggplot2)
vis_dat(Vaccin_DHB) + ggtitle("Missing Value Distribution")
```

There is no missing data in the dataset.

## Novelties

Outliers by DHB - First Dose
There are some high outliers are showing for some DHB. Extremely high outliers can be seen for Auckland Metro DHB. 
```{r outliers_DHB_firstDose}
#coef <- 1.5
#limits <- boxplot.stats(x = Vaccin_DHB$`First dose uptake per 1000 people`, coef = coef)$stats

#Vaccin_DHB$label <- ifelse(Vaccin_DHB$`First dose uptake per 1000 people` < limits[1] | Vaccin_DHB$`First dose uptake per 1000 people` > limits[5], rownames(Vaccin_DHB), NA)

ggplot(data = Vaccin_DHB, aes(x=`DHB of residence`, y=`First dose uptake per 1000 people`)) + 
geom_boxplot(aes(fill=`DHB of residence`), outlier.colour = "red")+ 
#ggrepel::geom_label_repel() +
labs(title = paste("Boxplots: First dose uptake per 1000 people"))+
theme(axis.title.y = element_blank(), axis.text.y = element_text(face = "bold"), axis.ticks.y = element_blank(), axis.text.x = element_text(face = "bold",angle = 90))+
theme(legend.position="down")
```

Outliers by DHB - Second Dose
There are some high outliers are showing for some DHB but the total number the DHB which are showing the outliers are less. Extremely high outliers can be seen for the same DHB -  Auckland Metro DHB. 
```{r outliers_DHB_secondDose}
ggplot(data = Vaccin_DHB, aes(x=`DHB of residence`, y=`Second dose uptake per 1000 people`)) + 
geom_boxplot(aes(fill=`DHB of residence`), outlier.colour = "red")+ 
labs(title = paste("Boxplots: Second dose uptake per 1000 people"))+
theme(axis.title.y = element_blank(), axis.text.y = element_text(face = "bold"), axis.ticks.y = element_blank(), axis.text.x = element_text(face = "bold",angle = 90))+
theme(legend.position="down")
```



Outliers by Age Group - First Dose
There are some high outliers are showing for the age group from 12 to 44. Low and hight outliers are showing for the age group 45-90+. Extremely igh oulier is showing for the age group from 25 to 29. 

```{r outliers_age_firstDose}
library(ggrepel)
ggplot(data = Vaccin_DHB, aes(x=`Age group`, y=`First dose uptake per 1000 people`)) + 
geom_boxplot(aes(fill=`Age group`), outlier.colour = "red")+ 
labs(title = paste("Boxplots: First dose uptake per 1000 people"))+
theme(axis.title.y = element_blank(), axis.text.y = element_text(face = "bold"), axis.ticks.y = element_blank(), axis.text.x = element_text(face = "bold",angle = 90))+
theme(legend.position="down")
```

Outliers by Age Group - Second Dose
There is not a big difference between outliers for the first and second dose by the Age Group. Basically the picture is pretty the same. 
```{r outliers_age_secondDose}
library(ggrepel)
ggplot(data = Vaccin_DHB, aes(x=`Age group`, y=`Second dose uptake per 1000 people`)) + 
geom_boxplot(aes(fill=`Age group`), outlier.colour = "red")+ 
labs(title = paste("Boxplots: Second dose uptake per 1000 people"))+
theme(axis.title.y = element_blank(), axis.text.y = element_text(face = "bold"), axis.ticks.y = element_blank(), axis.text.x = element_text(face = "bold",angle = 90))+
  theme(legend.position="down")
```

### DHB of Residence by ethnicity: Data overview

To simplify visualisation I have decided to reduce the Age group from 18 to 8 merging different Age Groups.

```{r plot1}
library(dplyr)
Vaccin_DHB<-Vaccin_DHB %>%
  mutate(Age_Group_min = case_when(
    grepl("12-15", as.character(`Age group`)) ~ "12-19",
    grepl("16-19", as.character(`Age group`)) ~ "12-19",
    grepl("20-24", as.character(`Age group`)) ~ "20-29",
    grepl("25-29", as.character(`Age group`)) ~ "20-29",
    grepl("30-34", as.character(`Age group`)) ~ "30-39",
    grepl("35-39", as.character(`Age group`)) ~ "30-39",
    grepl("40-44", as.character(`Age group`)) ~ "40-49",
    grepl("45-49", as.character(`Age group`)) ~ "40-49",
    grepl("50-54", as.character(`Age group`)) ~ "50-59",
    grepl("55-59", as.character(`Age group`)) ~ "50-59",
    grepl("60-64", as.character(`Age group`)) ~ "60-69",
    grepl("65-69", as.character(`Age group`)) ~ "60-69",
    grepl("70-74", as.character(`Age group`)) ~ "70+",
    grepl("75-79", as.character(`Age group`)) ~ "70+",
    grepl("80-84", as.character(`Age group`)) ~ "70+",
    grepl("85-89", as.character(`Age group`)) ~ "70+",
    grepl("90+", as.character(`Age group`)) ~ "70+",
    grepl("Various", as.character(`Age group`)) ~ "Various"
    ))

Vaccin_DHB$Age_Group_min<-factor(Vaccin_DHB$Age_Group_min)

library(ggplot2)
library(dplyr)

df1 = Vaccin_DHB%>%
  group_by(`DHB of residence`,`Age_Group_min`) %>%
  summarise(Mean = mean(`First dose uptake per 1000 people`), Sd = sd(`First dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`DHB of residence`))


ggplot(df1, aes(x = New_Var, y = Mean, fill = `Age_Group_min`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("First dose uptake per 1000 people by DHB for different Age Groups")+
  xlab("DHB of residence") +
  ylab("Number of vaccines admin (mean by Age group)")


```

```{r plot11}

df11 = Vaccin_DHB%>%
  group_by(`DHB of residence`,`Ethnic group`) %>%
  summarise(Mean = mean(`First dose uptake per 1000 people`), Sd = sd(`First dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`DHB of residence`))


ggplot(df11, aes(x = New_Var, y = Mean, fill = `Ethnic group`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("First dose uptake per 1000 people by DHB for different Ethnicity")+
  xlab("DHB of residence") +
  ylab("Number of vaccines admin (mean by Ethnic group)")


```

```{r plot2}

df2 = Vaccin_DHB%>%
  group_by(`DHB of residence`,`Age_Group_min`) %>%
  summarise(Mean = mean(`Second dose uptake per 1000 people`), Sd = sd(`Second dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`DHB of residence`))


ggplot(df2, aes(x = New_Var, y = Mean, fill = `Age_Group_min`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("Second dose uptake per 1000 people by DHB for different Age Group")+
  xlab("DHB of residence") +
  ylab("Number of vaccines admin (mean by Age group)")


```

```{r plot22}

df22 = Vaccin_DHB%>%
  group_by(`DHB of residence`,`Ethnic group`) %>%
  summarise(Mean = mean(`Second dose uptake per 1000 people`), Sd = sd(`Second dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`DHB of residence`))


ggplot(df22, aes(x = New_Var, y = Mean, fill = `Ethnic group`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("Second dose uptake per 1000 people by DHB for different Ethnicity")+
  xlab("DHB of residence") +
  ylab("Number of vaccines admin (mean by Ethnic group)")


```


```{r plot3}

df3 = Vaccin_DHB%>%
  group_by(`Ethnic group`,`Age_Group_min`) %>%
  summarise(Mean = mean(`First dose uptake per 1000 people`), Sd = sd(`First dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`Ethnic group`))


ggplot(df3, aes(x = New_Var, y = Mean, fill = `Age_Group_min`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("First dose uptake per 1000 people by Age group")+
  xlab("Ethnic group") +
  ylab("Number of vaccines admin (mean by Age group)")

```

```{r plot33}

df33 = Vaccin_DHB%>%
  group_by(`Ethnic group`,`Age_Group_min`) %>%
  summarise(Mean = mean(`Second dose uptake per 1000 people`), Sd = sd(`Second dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`Ethnic group`))


ggplot(df33, aes(x = New_Var, y = Mean, fill = `Age_Group_min`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("Second dose uptake per 1000 people by Age Group")+
  xlab("Ethnic group") +
  ylab("Number of vaccines admin (mean by Age group)")

```
```{r plot6}

df6 = Vaccin_DHB%>%
  group_by(`DHB of residence`,`Gender`) %>%
  summarise(Mean = mean(`First dose uptake per 1000 people`), Sd = sd(`First dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`DHB of residence`))


ggplot(df6, aes(x = New_Var, y = Mean, fill = `Gender`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("First dose uptake per 1000 people by DHB for different Gender")+
  xlab("DHB of residence") +
  ylab("Number of vaccines admin (mean by Gender)")

```

```{r plot66}

df66 = Vaccin_DHB%>%
  group_by(`DHB of residence`,`Gender`) %>%
  summarise(Mean = mean(`Second dose uptake per 1000 people`), Sd = sd(`Second dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`DHB of residence`))


ggplot(df66, aes(x = New_Var, y = Mean, fill =`Gender`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("Second dose uptake per 1000 people by DHB for different Gender")+
  xlab("DHB of residence") +
  ylab("Number of vaccines admin (mean by Gender)")

```
```{r plot7}

df7 = Vaccin_DHB%>%
  group_by(`Ethnic group`,`Gender`) %>%
  summarise(Mean = mean(`First dose uptake per 1000 people`), Sd = sd(`First dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`Ethnic group`))


ggplot(df7, aes(x = New_Var, y = Mean, fill = `Gender`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("First dose uptake per 1000 people by Ethnicity for different Gender")+
  xlab("Ethnic group") +
  ylab("Number of vaccines admin (mean by Gender)")

```

```{r plot77}

df77 = Vaccin_DHB%>%
  group_by(`Ethnic group`,`Gender`) %>%
  summarise(Mean = mean(`Second dose uptake per 1000 people`), Sd = sd(`Second dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`Ethnic group`))


ggplot(df77, aes(x = New_Var, y = Mean, fill =`Gender`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("Second dose uptake per 1000 people by Ethnicity for different Gender")+
  xlab("Ethnic group") +
  ylab("Number of vaccines admin (mean by Gender)")

```

```{r plot8}

df8 = Vaccin_DHB%>%
  group_by(`Age group`,`Gender`) %>%
  summarise(Mean = mean(`First dose uptake per 1000 people`), Sd = sd(`First dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`Age group`))


ggplot(df8, aes(x = New_Var, y = Mean, fill = `Gender`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("First dose uptake per 1000 people by Age Group for different Gender")+
  xlab("Age group") +
  ylab("Number of vaccines admin (mean by Gender)")

```

```{r plot88}

df88 = Vaccin_DHB%>%
  group_by(`Age group`,`Gender`) %>%
  summarise(Mean = mean(`Second dose uptake per 1000 people`), Sd = sd(`Second dose uptake per 1000 people`)) %>%
  mutate(New_Var = paste0(`Age group`))


ggplot(df88, aes(x = New_Var, y = Mean, fill =`Gender`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  ggtitle("Second dose uptake per 1000 people by Age Group for different Gender")+
  xlab("Age group") +
  ylab("Number of vaccines admin (mean by Age group)")

```

```{r plot4}
library(scales)
df4sum<-sum(Vaccin_DHB$`First dose administered`)
df4 = Vaccin_DHB%>%
  group_by(`Ethnic group`) %>%
  summarise(Sum = sum(`First dose administered`), Labels=percent(sum(`First dose administered`)/ df4sum ) )%>%
  mutate(New_Var = paste0(`Ethnic group`))

#remove values less than 1% as they are not significant
df4<-df4[which(df4$Labels>1),]

#plot a pie
pie(df4$Sum, labels = paste(df4$`Ethnic group`,df4$Labels), main="First Dose administered by Ethnicity for all DHB", col=terrain.colors(length(df4$Labels)))

```

```{r plot44}

df44sum<-sum(Vaccin_DHB$`Second dose administered`)
df44 = Vaccin_DHB%>%
  group_by(`Ethnic group`) %>%
  summarise(Sum = sum(`Second dose administered`), Labels=percent(sum(`Second dose administered`)/ df44sum ) )%>%
  mutate(New_Var = paste0(`Ethnic group`))

#remove values less than 1% as they are not significant
df44<-df44[which(df44$Labels>1),]

#plot a pie
pie(df44$Sum, labels = paste(df44$`Ethnic group`,df44$Labels), main="Second Dose administered by Ethnicity for all DHB", col=terrain.colors(length(df44$Labels)))
  
```

```{r plot4a}
library(scales)
df4a = Vaccin_DHB%>%
  group_by(`Ethnic group`) %>%
  summarise(Sum = sum(`First dose administered`), pop=sum(`Population`) )%>%
  mutate(New_Var = paste0(`Ethnic group`))

ggplot(df4a, aes(x = New_Var, y =Sum/pop, fill =`Ethnic group`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  scale_y_continuous(labels=scales::percent) +
  ggtitle("First dose dose vaccin admin in % of a total pupolation by Ethnicity for all DHB")+
  xlab("Ethnic Group") +
  ylab("First dose administered, %")

```

```{r plot44a}

library(scales)
df44a = Vaccin_DHB%>%
  group_by(`Ethnic group`) %>%
  summarise(Sum = sum(`Second dose administered`), pop=sum(`Population`) )%>%
  mutate(New_Var = paste0(`Ethnic group`))

ggplot(df44a, aes(x = New_Var, y =Sum/pop, fill =`Ethnic group`)) +
  geom_bar(stat = "identity", color = "black",position = position_dodge(), width = 0.7)+
  theme(axis.text.x = element_text(face = "bold",angle = 90),
  legend.title = element_blank()) +
  scale_y_continuous(labels=scales::percent) +
  ggtitle("Second dose vaccin admin in % of a total pupolation by Ethnicity for all DHB")+
  xlab("Ethnic Group") +
  ylab("Second dose administered, %")
  
```
```{r plot5}

df5sum<-sum(Vaccin_DHB$`First dose administered`)
df5 = Vaccin_DHB%>%
  group_by(`Age_Group_min`) %>%
  summarise(Sum = sum(`First dose administered`), Labels=percent(sum(`First dose administered`)/ df5sum ) )%>%
  mutate(New_Var = paste0(`Age_Group_min`))

#remove values less than 1% as they are not significant
df5<-df5[which(df5$Labels>1),]

#plot a pie
pie(df5$Sum, labels = paste(df5$Labels, "for Age Group ", df5$`Age_Group_min`), main="First Dose administered by Age Group for all DHB", col=terrain.colors(length(df5$Labels)))

```

```{r plot55}
df55sum<-sum(Vaccin_DHB$`Second dose administered`)
df55 = Vaccin_DHB%>%
  group_by(`Age_Group_min`) %>%
  summarise(Sum = sum(`Second dose administered`), Labels=percent(sum(`Second dose administered`)/ df55sum ) )%>%
  mutate(New_Var = paste0(`Age_Group_min`))

#remove values less than 1% as they are not significant
df55<-df55[which(df55$Labels>1),]

#plot a pie
pie(df55$Sum, labels = paste(df55$Labels,"for Age Group",df55$`Age_Group_min`), main="Second Dose administered by Age Group for all DHB", col=terrain.colors(length(df55$Labels)))
  
```

```{r plot9}

df9sum<-sum(Vaccin_DHB$`First dose administered`)
df9 = Vaccin_DHB%>%
  group_by(`Gender`) %>%
  summarise(Sum = sum(`First dose administered`), Labels=percent(sum(`First dose administered`)/ df9sum ) )%>%
  mutate(New_Var = paste0(`Gender`))

#remove values less than 1% as they are not significant
df9<-df9[which(df9$Labels>1),]

#plot a pie
pie(df9$Sum, labels = paste(df9$Labels, df9$`Gender`), main="First Dose administered by Gender for all DHB", col=terrain.colors(length(df9$Labels)))

```

```{r plot99}
df99sum<-sum(Vaccin_DHB$`Second dose administered`)
df99 = Vaccin_DHB%>%
  group_by(`Gender`) %>%
  summarise(Sum = sum(`Second dose administered`), Labels=percent(sum(`Second dose administered`)/ df99sum ) )%>%
  mutate(New_Var = paste0(`Gender`))

#remove values less than 1% as they are not significant
df99<-df99[which(df99$Labels>1),]

#plot a pie
pie(df99$Sum, labels = paste(df99$Labels,df99$`Gender`), main="Second Dose administered by Gender for all DHB", col=terrain.colors(length(df99$Labels)))
  
```

